---
title: 'Statistical report: Anemia in patients with heart failure managed in primary care'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.2.6. 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = FALSE)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1}. All-cause mortality was included as the Nelson-Aalen estimator.

## Baseline characteristics

```{r, child = "../src/tabprev.Rmd"}

```

```{r, child = "../src/tab1.Rmd"}

```

```{r, child = "../src/tabcentre.Rmd"}

```

\clearpage

## Associations with anemia

Associations between anemia and baseline characteristics were evaluated using a multivariable logistic 
regression with variables indicated in Table \ref{tab:tab1}. 
The variables were selected based on clinical relevance. 

```{r, child = "../src/predictors.Rmd"}

```

### Assumptions

Possible outliers were inspected with Cook's distance and multicollinearity with the Variance Inflation factor.
=> No action deemed necessary.

\clearpage

## Association between anemia and outcomes

The following outcomes are considered: 

- All-cause mortality
- First hospitalization

Data were censored at 6 years of follow-up, 2019-12-31 or death/emigration. 

The outcomes were presented with cumulative incidence curves. 
Cox proportional hazards regressions were 
used to model the time to event, partly crude and partly adjusted for variables indicated in 
Table \ref{tab:tab1}. The variables were selected based on clinical relevance.

As a consistency analysis all-cause hospitalization was modelled using a sub-distributional hazards model [@fg] 
where death was treated as a competing event. 

The association between anemia and the outcomes in selected subgroups were 
investigated using Cox proportional hazards regressions were, together with 
the variables indicated in Table \ref{tab:tab1}, the interaction between anemia 
and the respective variable was included, thereby giving the interpretation of a 
subgroup analysis but with the additional advantage of also being able to test 
for differences in effect between subgroups. 

The median (min-max) follow-up is 
`r rsdata %>% summarise(med = fn(median(sos_outtime_death6y / 365.25), dig = 1),
                                             min = fn(min(sos_outtime_death6y / 365.25), dig = 1),
                                             max = fn(max(sos_outtime_death6y / 365.25), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r rsdata %>% summarise(sumpy = fn(sum(sos_outtime_death6y) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

```{r, child = "../src/outforestsub.Rmd"}

```

\clearpage

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the outcomes. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> No action deemed necessary. 

### Residual confounding

In order to assess how robust the association between anemia and the outcomes are 
to residual confounding the E-value 
was calculated [@evalue]. The E-value is interpreted as the minimum Risk Ratio 
that an unmeasured confounder would need to have with both anemia and 
the outcome, conditional on the variables currently adjusted 
for in the model, to fully eliminate the association between anemia and 
outcome. See Table \ref{tab:evalueprint2} in order to compare the magnitude of the E-value 
within the context of the study [@evalue2]. 

```{r, child = "../src/evalue.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/anemiapc. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
