---
title: 'Statistical report: Anemia in patients with heartfailure managed in primary care'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.2.4. 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart", scale_down = FALSE)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1}. All-casuse mortality was included as the Nelson-Aalen estimator.

## Baseline characteristics

```{r, child = "../src/tabprev.Rmd"}

```

```{r, child = "../src/tab1.Rmd"}

```

\clearpage

## Associations with anemia

Associations between anemia and baseline characteristics were evaluated using a multivariable logistic 
regression with variables indicated in Table \ref{tab:tab1}. 
The variables were selected based on clinical relevance. 

```{r, child = "../src/predictors.Rmd"}

```

\clearpage

## Association between anemia and outcomes

The following outcomes are considered: 

- All-cause mortality
- First hospitalization

Data were censored after 6 years, at 2019-12-31 or death/emigration. 

All-cause mortality was presented with a Kaplan-Meier curve and first 
hospitalization with a cumulative incidence curve. 
Cox proportional hazards regressions were 
used to model the time to event, partly crude and partly adjusted for variables indicated in 
Table \ref{tab:tab1}. The variables were selected based on clinical relevance.

As a consistency analysis all-cause hospitalization was modelled using a sub-distributional hazards model [@fg] 
were death was treated as a competing event. 

The association between anemia and the outcomes in selected subgroups were 
investigated using Cox proportional hazards regressions were, together with 
the variables indicated in Table \ref{tab:tab1}, the interaction between anemia 
and the respective variable was included, thereby giving the interpretation of a 
subgroup analysis but with the additional advantage of also being able to test 
for differences in effect between subgroups. 

The median (min-max) follow-up is 
`r rsdata %>% summarise(med = fn(median(sos_outtime_death6y / 365.25), dig = 1),
                                             min = fn(min(sos_outtime_death6y / 365.25), dig = 1),
                                             max = fn(max(sos_outtime_death6y / 365.25), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` years for a total of 
                                   `r rsdata %>% summarise(sumpy = fn(sum(sos_outtime_death6y) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.


```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

```{r, child = "../src/outforestsub.Rmd"}

```

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. 
Possible outliers were visually inspected by plotting the dfbetas. 
=> XXXX. 

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/anemiapc. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
